---
title: "Homework 3"
subtitle: "ECON 470, Spring 2025"
author: "Ethan Murakami"
format:
  pdf:
    output-file: "murakami_e_hmwk2_s1"
    output-ext:  "pdf"
    header-includes:
      - \usepackage{float}
      - \floatplacement{table}{H}
---

```{r}
#| include: false

if (!require("pacman")) install.packages("pacman")
pacman::p_load(tidyverse, ggplot2, dplyr, lubridate, readr, readxl, hrbrthemes, fixest,
               scales, gganimate, gapminder, gifski, png, tufte, plotly, OECD,
               ggrepel, survey, foreign, devtools, pdftools, kableExtra, modelsummary,
               kableExtra)
```

Here is a link to my repository: {https://github.com/bemur3/hmwk3}

\newpage 

```{r}
#| include: false
#| eval: true


load("/Users/ethanmurakami/Documents/GitHub/hmwk3/submission2/Hwk3_workspace.Rdata")
```


\newpage 

## 1. Present a bar graph showing the proportion of states with a change in their cigarette tax in each year from 1970 to 1985.

```{r} 
#| echo: false
#| label: tax.changeplot

print(tax.change.plot)
```


\newpage 

## 2. 

```{r} 
#| echo: false
#| label: avtgtax.priceplot

print(avtgtax.price.plot)
```

\newpage

## 3. 

```{r} 
#| echo: false
#| label: top.statesplot

print(top.states.plot)
```

\newpage

## 4. 

```{r} 
#| echo: false
#| label: bottom.statesplot

print(bottom.states.plot)
```

\newpage

## 5. 

```{r} 
#| echo: false
#| label: mergedstatesplot

print(merged.states.plot)
```

\newpage

## 6-10 Made summary table but won't load into quarto pdf. putting code below, please view table in submission history on grade scope

## Load libraries
library(modelsummary)
library(gt)

# Create a custom coefficient mapping
coef_map <- c(
  "log_price" = "Log Price",
  "log_price_b" = "Log Price",
  "fit_log_price" = "Log Price",  # Instrumented log price for IV model
  "fit_log_price_b" = "Log Price",  # Instrumented log price for IV model (1991-2015)
  "log_tax" = "Total Tax",
  "log_tax_b" = "Total Tax"
)

# Define the list of models
models <- list(
  "OLS" = elasticity_model,
  "IV" = iv_model,
  "First Stage" = first_stage,
  "Reduced Form" = reduced_form,
  "OLS" = elasticity_model_b,
  "IV" = iv_model_b,
  "First Stage" = first_stage_b,
  "Reduced Form" = reduced_form_b
)

# Generate the regression table 
modelsummary(models,
  fmt = 3,  # Round to 3 decimal places
  statistic = "({std.error})",  # Show standard errors in parentheses
  stars = TRUE,  # Add significance stars
  coef_map = coef_map,  # Use the corrected coefficient mapping
  coef_omit = "Intercept",  # Remove intercept
  gof_omit = "IC|Adj|Log|RMSE",  # Remove extra goodness-of-fit statistics
  output = "gt"
) %>%
  tab_header(
    title = "Regression Estimates (OLS & IV)",
    subtitle = "Comparison of 1970-1990 and 1991-2015"
  ) %>%
  tab_spanner(label = "1970-1990 Models", columns = 2:5) %>%
  tab_spanner(label = "1991-2015 Models", columns = 6:9) %>%
  tab_options(table.font.size = "medium")



